\documentclass{article}
\usepackage{ee105}

%% Makes figure labels bold
\usepackage[labelfont=bf]{caption}

\begin{document}

%% Prevent headings on the first page, since we're not using \maketitle
\thispagestyle{plain}

%% Standard header. Title of the lab goes here.
\report{8}{Multi-stage Amplifiers}
\name

\begin{enumerate}

	\item[3.1.3] Measure $I_{BIAS}$ and the DC voltage at $v_{OUT}$.

	\begin{align*}
    		\boxed{I_{BIAS} = ~~~~~~~~~~~~~~~~~~~~ } \\
    		\boxed{v_{OUT} = ~~~~~~~~~~~~~~~~~~~~ } \\
  	\end{align*}

	\item[3.1.4] Using the oscilloscope, plot both the input $v_{IN}$ and the output $v_{OUT}$. Sketch the waveforms you observe.

\begin{figure}[!htb]
\begin{center}
\begin{asy}
import math;
size(250,0);

add(shift(-5,-4)*grid(10,8));

dot((0,0),black);
\end{asy}
\end{center}
\end{figure}
	\item[3.1.5] Why is $v_{OUT}$ not sinusoidal? \\~\\~\\~\\

	\item[3.1.7] What is the peak-to-peak voltage of the output waveform (at $v_L$) with the load resistor? What is the gain of the amplifier with the resistive load? \\

	\begin{align*}
		\boxed{v_{L,p-p} = ~~~~~~~~~~~~~~~~~~~} \\
    		\boxed{\left|\frac{v_L}{v_{in}}\right| = ~~~~~~~~~~~~~~~~~~~ } \\
  	\end{align*}

	\item[3.2.1] Can you hear anything when the speaker is directly hooked up to the function generator? 

	\begin{align*}
		\boxed{\text{Yes or No}} \\
	\end{align*}

	\item[3.2.3] Can you hear anything when the speaker is hooked up to the output of the amplifier? 

	\begin{align*}
		\boxed{\text{Yes or No}} \\
	\end{align*}

	\item[3.2.4] Measure $I_{BIAS1}$, $I_{BIAS2}$, and the DC voltages at $v_{OUT1}$ and $v_{OUT2}$.

	\begin{align*}
    		\boxed{I_{BIAS1} = ~~~~~~~~~~~~~~~~~~~ } \\
		\boxed{I_{BIAS2} = ~~~~~~~~~~~~~~~~~~~ } \\
    		\boxed{v_{OUT1,DC} = ~~~~~~~~~~~~~~~~~~~ } \\
		\boxed{v_{OUT2,DC} = ~~~~~~~~~~~~~~~~~~~ } \\
  	\end{align*}

	\item[3.2.5] Measure $V_{BE}$ of $Q_2$. Is the DC voltage at $v_{OUT1}$ enough to bias $Q_2$ in the forward active region?

	\begin{align*}
		\boxed{V_{BE} = ~~~~~~~~~~~~~~~~~~~~~ } \\
		\boxed{\text{Yes or No} } \\
	\end{align*}

	\item[3.2.6] Using the oscilloscope, plot both the input $v_{IN}$ and the output $v_{OUT2}$. Sketch these waveforms (a plot is on the following page).

\begin{figure}[!htb]
\begin{center}
\begin{asy}
import math;
size(250,0);

add(shift(-5,-4)*grid(10,8));

dot((0,0),black);
\end{asy}
\end{center}
\end{figure}
	\item[3.2.7] Measure the gain $v_{out2}/v_{in}$.

	\begin{align*}
    		\boxed{\left|\frac{v_{out2}}{v_{in}}\right| = ~~~~~~~~~~~~~~~~~~~~ } \\
  	\end{align*}

	\item[3.2.8] Now increase the DC offset of the input waveform to \unit{620}{\milli\volt}. What happens to the waveform at $v_{OUT2}$?

\end{enumerate}

\end{document}
